---
title: "Reproducible Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and Preprocessing the Data
``` {R}
data <- read.csv("activity.csv")
```
### What is mean total number of steps taken per day?

``` {R}
stepsPerDay <- aggregate(steps ~date, data, sum)
hist(stepsPerDay$steps, main=paste("Total steps Each Day"), col="red" , xlab="Number of steps")
rmean <- mean(stepsPerDay$steps)
rmedian <- median(stepsPerDay$steps)
```
The `mean` is `r rmean` and the median is `r rmedian`



## What is the average daily activity pattern?

```{r}
stepsByInterval <- aggregate(steps ~ interval, data, mean)

plot(stepsByInterval$interval,stepsByInterval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
maxInterval <- stepsByInterval[which.max(stepsByInterval$steps), 1]
```
The 5-minute interval on average across all the days in the data set containing the maximum number of steps is `r maxInterval`.

## Computing missing values
#### Total number of missing values
```{r}
incomplete <- sum(!complete.cases(data))
```

The total number of missing values in the dataset
is `r incomplete`
```{r}
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), stepsByInterval$steps[match(data$interval, stepsByInterval$interval)], data$steps))
```
Missing values were imputed by inserting the average for each interval.
```{R}
imputed_data[as.character(imputed_data$date) == "2012-10-01", 1] <- 0
```
NAs then were assumed to be zeros to fit the rising trend of the data.

Recount total steps by day and create Histogram.
```{R}
stepsPerDayi <- aggregate(steps ~ date, imputed_data, sum)
hist(stepsPerDayi$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps")
stepsPerDay <- aggregate(steps ~date, data, sum)
hist(stepsPerDay$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```



##Calculate new mean for imputed data.
```{R}
rmean.i <- mean(stepsPerDayi$steps)
rmedian.i <- median(stepsPerDayi$steps)
```

Calculate difference between imputed and non-imputed data.
```{R}
mean_diff <- rmean.i - rmean
med_diff <- rmedian.i - rmedian
```


Calculate total difference.
```{R}
total_diff <- sum(stepsPerDayi$steps) - sum(stepsPerDay$steps)
```

The imputed data mean is `r rmean.i`
The difference between the non-imputed mean and imputed mean is `r mean_diff`

The difference between total number of steps between imputed and non-imputed data is r total_diff. Thus,there were `r total_diff` more steps in the imputed data.

##Are there differences in activity patterns between weekdays and weekends?
```{R}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```
